Geometric Filters for Metric Spaces

### 2018-11-24 paper/draft.tex
Interim

In this section we consider the standard problem of finding an approximate nearest neighbor: given
$n$ points $P = \Set{p_1, \dots, p_n}$, a query point $q$ and a real number $\eps > 0$,
find $p_i$ such that $\dist(p_i, q) \leq (1 + \eps) \min_{k} \dist(q, p_k)$. We assume for simplicity
that all exact pairwise distances $\dist(p_i, p_j)$ are already computed. Our goal is to reduce the number
of computed distances $d(p_i, q)$.

@@

In order to prove this theorem, we can think of the algorithm as

\begin{lemma}
\end{lemma}

>>>>

In this section we consider the standard problem of finding an approximate nearest neighbor: given
$n$ points $P = \Set{p_1, \dots, p_n}$, a query point $q$ and a real number $\eps > 0$,
find $p_i$ such that $\dist(p_i, q) \leq (1 + \eps) \min_{k} \dist(q, p_k)$. This notation will be fixed throughout this 
section, and we shall also use the shorthand notation
\[
    r_i := \dist(p_i, q).
\]
We assume for simplicity
that all exact pairwise distances $\dist(p_i, p_j)$ are already computed. Our goal is to reduce the number
of computed distances $d(p_i, q)$.

@@

The proof is based on the following lemma.
\begin{lemma}
    There exists a constant $C$ that depends only on $\eps$ and the doubling dimension
    such that after computing at most $C$ distances $\dist(p_k, q)$
    either the algorithm verifies that the current candidate $p_i$ is a correct answer and terminates
    or the algorithm finds another candidate $p_j$ such that $r_j < r_i$.
\end{lemma}
\begin{proof}
    First, all points that are not contained in the ball $B(p_i, 2r_i)$
    are discarded without computing the distance to them.
    Now we have to examine all points $p_k$ inside this ball. 
    There are three possibilities for $r_k$:
    \begin{enumerate}
        \item $r_i < r_1 / (1 + \eps)$. This outcome refutes the statement $p_1$ is an approximate
            nearest neighbor and we can stop verification.

    Second, after we computed the distance $\dist(p_i, q)$
    and checked that it is greater than $r_1 / (1 + \eps)$, we can discard all points inside
    the ball $B(p_1, r_i - r_1 / (1+ \eps))$
\end{proof}

First we consider a simplified version of the problem, where the candidate for being an approximate nearest neighbor
is given, and we only have to verify this. Without loss of generality, assume that the candidate is $p_1$.
